---
title: "GAM on affordability"
output:
  pdf_document: default
  html_document: default
date: "2023-07-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
rm(list=ls())

###Setting the working directory as the project directory
setwd("~/Documents/GitHub/Nonparametric-Statistics-project/src/Markdowns")

inputpath = "../../data"
outputpath = "../../data"

#Imports
library(tidyr)
library(dplyr)
library(lme4)
library(pbapply)
library(mgcv)
library(conformalInference)
library(ggplot2)
library(progress)
library(parallel)
library(pbapply)
#install.packages("qgam")
library(qgam)


seed=2022
B=1000
```


```{r, include=FALSE}
####Permutational test diagnostics
diagnostic_permutation <- function(T20, T2) {
  B <- length(T2)
  # Compare real test statistic with the ones given by the permuted data
  hist(T2,breaks = 100, xlim = range(c(T2, T20)))
  abline(v = T20, col = 3, lwd = 4)
  # Empirical cumulative distribution function
  plot(ecdf(T2))
  abline(v = T20, col = 3, lwd = 4)
  # P-value
  p_val <- sum(T2 >= T20) / B
  cat("p-value: ", p_val)
}

diagnostic_bootstrap <- function(distro, obs, alpha) {  #distro = T.boot, 
                                                        #obs = predict(fit,x)$y, 
                                                        #      or T.obs
                                                        #x potrebbe essere un nuovo punto
    variance_pred <- var(distro)
    print(paste("Variance: ", variance_pred))
    sd_pred <- sd(distro)
    print(paste("Standard deviation: ", sd_pred))
    bias_pred <- mean(distro) - obs
    print(paste("Bias: ", bias_pred))
    MSE_pred <- variance_pred + bias_pred^2
    print(paste("MSE: ", MSE_pred))
    # computing quantiles
    right.quantile <- quantile(distro, 1 - alpha / 2)
    left.quantile <- quantile(distro, alpha / 2)
    # REVERSE!!!!-percentile ----- guardare come si fanno i non reverse!!!!!credo invertendo i due
    CI <- c(
        obs - (right.quantile - obs),
        obs,
        obs - (left.quantile - obs)
    )
    names(CI) <- c("lower", "center", "upper")
    print(CI)
    plot(ecdf(distro), main = "Bootstrap distribution")
    abline(v = CI[2], col = 3, lwd = 2)
    abline(v = CI[c(1, 3)], lty = 3)
}
#Import data
data<-read.table("../../data/final_dataset_2007-2020.txt",
                 header=T, 
                 sep='',
                 check.names = FALSE)
data$Year<-as.numeric(data$Year)

#creating unified mpower indes
#Creating a unified index
min_index=12 # since 1 is not possible
max_index=29 #6*5 minus 1 for monitor which is up to 5

data$mpower_all=((data$Taxes+
                    data$Help+
                    data$Warn+
                    data$Bans+
                    data$Protect+
                    data$Monitor)-min_index)/(max_index-min_index)
boxplot(data$mpower_all~data$Country)
boxplot(data$mpower_all~as.factor(data$Year))
#Seems that the countries have increased their policies across years!
data$Year<-as.numeric(data$Year)

data_gam<-data
data_gam=data_gam[data_gam$Year!='2007',]
data_gam=data_gam[data_gam$Year!='2008',]

#I also remove japan as it has no education data
data_gam=data_gam[data_gam$Country!='Japan',]

#data_gam$Campaigns<-as.numeric(data_gam$Campaigns)
data_gam$Bans<-as.numeric(data_gam$Bans)
data_gam$Monitor<-as.numeric(data_gam$Monitor)
data_gam$Help<-as.numeric(data_gam$Help)
data_gam$Protect<-as.numeric(data_gam$Protect)
data_gam$Warn<-as.numeric(data_gam$Warn)
data_gam$Taxes<-as.numeric(data_gam$Taxes)
data_gam$GDP<-log(data_gam$GDP)

```

I start by running the model for males and females


```{r, }
model_gam_both <- gam(
  Prevalence_both ~
    Year +
    Country+
    s(HDI, bs = 'cr') +
    s(GDP, bs = 'cr') +
    s(Education, bs = 'cr') +  
   Affordability,
  data = data_gam
)
#summary(model_gam_both)
#plot(model_gam_both)
plot(model_gam_both$residuals)
```

The countries clearly explain the majority of the variance in the data
Let's see if first we can remove the GDP from the model
H0:GDP term is 0, vs H1 is different than zero


```{r, }
summ<-summary(model_gam_both)
T0<-abs(summ$s.table[2,3])
T0

model_gam_noGDP<-gam(
  Prevalence_both ~
    Year +
    Country+
    s(HDI, bs = 'cr') +
    s(Education, bs = 'cr') +  
    Affordability,
  data = data_gam
)
#summary(model_gam_noGDP)
  
res<-model_gam_noGDP$residuals
fitted.values<-model_gam_noGDP$fitted.values

set.seed(seed)
T2<-numeric(B)
n<-nrow(data_gam)
for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res[permutation]
  response.perm <- fitted.values + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      s(data_gam$GDP, bs = 'cr') +
      s(data_gam$Education, bs = 'cr') +  
      data_gam$Affordability
  )
  T2[perm] <- abs(summary(model.perm)$s.table[2,3])
}
diagnostic_permutation(T0,T2)
```

We cannot reject the null hypothesis that the GDP is 0

```{r, }
model_gam_noGDPnoEDU<-gam(
  Prevalence_both ~
    Year +
    Country+
    s(HDI, bs = 'cr')+
    Affordability,
  data=data_gam
)

res<-model_gam_noGDPnoEDU$residuals
fitted.values<-model_gam_noGDPnoEDU$fitted.values

T0<-summary(model_gam_noGDP)$s.table[2,3]
T0

set.seed(seed)
T2<-numeric(B)
n<-nrow(data_gam)
for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res[permutation]
  response.perm <- fitted.values + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      s(data_gam$Education, bs = 'cr') +  
      data_gam$Affordability
  )
  T2[perm] <- abs(summary(model.perm)$s.table[2,3])
}
diagnostic_permutation(T0,T2)
```

We reject the null hypothesis and keep education

```{r, }
model_gam_noGDPnoHDI<-gam(
  Prevalence_both ~
    Year +
    Country+
    s(Education, bs = 'cr')+
    Affordability,
  data=data_gam
)

summ<-summary(model_gam_noGDP)
T0<-summ$s.table[1,3]
T0


res<-model_gam_noGDPnoHDI$residuals
fitted.values<-model_gam_noGDPnoHDI$fitted.values

set.seed(seed)
T2<-numeric(B)
n<-nrow(data_gam)
for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res[permutation]
  response.perm <- fitted.values + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      s(data_gam$Education, bs = 'cr') +  
      data_gam$Affordability
  )
  T2[perm] <- abs(summary(model.perm)$s.table[1,3])
}
diagnostic_permutation(T0,T2)
```

#We cannot reject H0 and we remove HDI, and I move to the affordability

```{r, }
model.gam.noaffordability<-gam(
  Prevalence_both ~
    Year +
    Country+
    s(Education, bs = 'cr'),
  data=data_gam
)
#summary(model.gam.noaffordability)
plot(model.gam.noaffordability)
T0<-abs(summary(model_gam_noGDP)$p.table[39,3])
T0

res<-model.gam.noaffordability$residuals
fitted.values<-model.gam.noaffordability$fitted.values

set.seed(seed)
T2<-numeric(B)
n<-nrow(data_gam)

for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res[permutation]
  response.perm <- fitted.values + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      s(data_gam$Education, bs = 'cr') +  
      data_gam$Affordability
  )
  T2[perm] <- abs(summary(model.perm)$p.table[39,3])
}
diagnostic_permutation(T0,T2)
```

It is not significant at alpha=0.1, I can remove it
At the net of the analysis it seems that the affordability
significantly and negatively on prevalence

```{r, }
model.final.both<-gam(
  data_gam$Prevalence_both ~
    data_gam$Year +
    data_gam$Country+
    s(data_gam$Education, bs = 'cr')  
)
#summary(model.final.both)
```

#############################
#############################
# Now I try on males
#############################
#############################

```{r, }
model_gam_males <- gam(
  Prevalence_males ~
    Year +
    Country+
    s(HDI, bs = 'cr') +
    s(GDP, bs = 'cr') +
    s(Education_males, bs = 'cr') +  
    Affordability,
  data = data_gam
)
#summary(model_gam_males)
#plot(model_gam_males)

summ_males<-summary(model_gam_males)
T0.males<-abs(summ_males$s.table[2,3])
T0.males


model_gam_noGDP.males<-gam(
  Prevalence_males ~
    Year +
    Country+
    s(HDI, bs = 'cr') +
    s(Education_males, bs = 'cr') +  
    Affordability,
  data = data_gam
)
#summary(model_gam_noGDP.males)

res.males<-model_gam_noGDP.males$residuals
fitted.values.males<-model_gam_noGDP.males$fitted.values


T2.males<-numeric(B)
n<-nrow(data_gam)
set.seed(seed)
for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res.males[permutation]
  response.perm <- fitted.values.males + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      s(data_gam$GDP, bs = 'cr') +
      s(data_gam$Education_males, bs = 'cr') +  
      data_gam$Affordability
  )
  T2.males[perm] <- abs(summary(model.perm)$s.table[2,3])
}
diagnostic_permutation(T0.males,T2.males)
```

We cannot reject H0, hence we can simplify the model and remove GDP

Now we try to remove the HDI variable

```{r, }
model.gam.noHDI<-gam(
  Prevalence_males ~
    Year +
    Country+
    s(Education_males, bs = 'cr')+
    Affordability,  
  data = data_gam
)
#summary(model.gam.noHDI)
T0<-summary(model_gam_noGDP.males)$s.table[1,3]
T0

res<-model.gam.noHDI$residuals
fitted.values<-model.gam.noHDI$fitted.values

set.seed(seed)
T2<-numeric(B)
n<-nrow(data_gam)
for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res[permutation]
  response.perm <- fitted.values + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      s(data_gam$Education_males, bs = 'cr') +  
      data_gam$Affordability
  )
  T2[perm] <- abs(summary(model.perm)$s.table[1,3])
}
diagnostic_permutation(T0,T2)
```

We cannot reject H0, we can remove HDI.

I now try to remove education

```{r, }
model.gam.noedu.males<-gam(
  Prevalence_males ~
    Year +
    Country+
    Affordability,  
  data = data_gam
)
#summary(model.gam.noedu.males)

T0<-abs(summary(model.gam.noHDI)$s.table[1,3])
T0

res<-model.gam.noedu.males$residuals
fitted.values<-model.gam.noedu.males$fitted.values

set.seed(seed)
T2<-numeric(B)
n<-nrow(data_gam)
for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res[permutation]
  response.perm <- fitted.values + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$Education_males, bs = 'cr') +  
      data_gam$Affordability
  )
  T2[perm] <- abs(summary(model.perm)$s.table[1,3])
}
diagnostic_permutation(T0,T2)
```

We cannot reduce the model and we should keep the education. Now we try to remove affordability
```{r, }
model.gam.noaffordability<-gam(
  Prevalence_males ~
    Year +
    Country+
    s(Education_males, bs = 'cr'),  
  data = data_gam
)
#summary(model.gam.noaffordability)
T0<-abs(summary(model.gam.noHDI)$p.table[39,3])
T0

res.males<-model.gam.noaffordability$residuals
fitted.values.males<-model.gam.noaffordability$fitted.values


T2<-numeric(B)
n<-nrow(data_gam)

for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res.males[permutation]
  response.perm <- fitted.values.males + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$Education_males, bs = 'cr') +  
      data_gam$Affordability
  )
  T2[perm] <- abs(summary(model.perm)$p.table[39,3])
}
diagnostic_permutation(T0,T2)
```

So we can remove affordability
So we have that we cannot say affordability impacts on males




#############################
#############################
# On females
#############################
#############################

We start by checking if we can remove GDP

```{r, }
model_gam_females1 <- gam(
  Prevalence_females ~
    Year +
    Country+
    s(HDI, bs = 'cr') +
    s(GDP, bs = 'cr') +
    s(Education_females, bs = 'cr') +  
    Affordability,
    data = data_gam
)
#summary(model_gam_females1)
#plot(model_gam_females1)


# Compute the distributions
summ_females<-summary(model_gam_females1)
T0.females<-summ_females$s.table[2,3]
T0.females

model_gam_noGDP.females<-gam(
  Prevalence_females ~
    Year +
    Country+
    s(HDI, bs = 'cr') +
    s(Education_females, bs = 'cr') +  
    Affordability,
    data = data_gam
)
#summary(model_gam_noGDP.females)

res.females<-model_gam_noGDP.females$residuals
fitted.values.females<-model_gam_noGDP.females$fitted.values


T2.females<-numeric(B)
n<-nrow(data_gam)

for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res.females[permutation]
  response.perm <- fitted.values.females + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      s(data_gam$GDP, bs = 'cr') +
      s(data_gam$Education_females, bs = 'cr') +  
      data_gam$Affordability
  )
  T2.females[perm] <- abs(summary(model.perm)$s.table[2,3])
}
diagnostic_permutation(T0.females,T2.females)
```

We cannot reject H0, hence we can simplify the model, so we remove the GDP.
Now we try to remove education

```{r, }
model_gam_females.noedu <- gam(
  Prevalence_females ~
    Year +
    Country+
    s(HDI, bs = 'cr') +
    Affordability,  
  data = data_gam
)
#summary(model_gam_females.noedu)
summ_females<-summary(model_gam_noGDP.females)

T0.females<-abs(summ_females$s.table[2,3])
T0.females

res.females<-model_gam_females.noedu$residuals
fitted.values.females<-model_gam_females.noedu$fitted.values

T2.females<-numeric(B)
n<-nrow(data_gam)

for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res.females[permutation]
  response.perm <- fitted.values.females + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      s(data_gam$Education_females, bs = 'cr') +  
      data_gam$Affordability
  )
  T2.females[perm] <- abs(summary(model.perm)$s.table[2,3])
}
diagnostic_permutation(T0,T2.females)
```

So we cannot reject H0, we remove education.
Now we try to remove HDI



```{r, }
model_gam_females.nohdi <- gam(
  Prevalence_females ~
    Year +
    Country+
    Affordability,
  data = data_gam
)
#summary(model_gam_females.nohdi)

T0<-summary(model_gam_females.noedu)$s.table[1,3]
T0

res.females<-model_gam_females.nohdi$residuals
fitted.values.females<-model_gam_females.nohdi$fitted.values

T2.females<-numeric(B)
n<-nrow(data_gam)
set.seed(seed)
for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res.females[permutation]
  response.perm <- fitted.values.females + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      data_gam$Affordability
  )
  T2.females[perm] <- abs(summary(model.perm)$s.table[1,3])
}
diagnostic_permutation(T0,T2.females)
```

We reject H0, cannot remove HDI

Now we try to remove affordability

```{r, }
model_gam_females.noaffordability <- gam(
  Prevalence_females ~
    Year +
    Country+
    s(HDI, bs = 'cr'),
  data = data_gam
)
#summary(model_gam_females.noaffordability)
#plot(model_gam_females.noaffordability)

summ_females<-summary(model_gam_females.noedu)
T0.females<-abs(summ_females$p.table[39,3])
T0.females

res.females<-model_gam_females.noaffordability$residuals
fitted.values.females<-model_gam_females.noaffordability$fitted.values

T2.females<-numeric(B)
n<-nrow(data_gam)
set.seed(seed)
for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res.females[permutation]
  response.perm <- fitted.values.females + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      data_gam$Affordability
  )
  T2.females[perm] <- abs(summary(model.perm)$p.table[39,3])
}

diagnostic_permutation(T0.females,T2.females)
```

 I can reject H0 and I keep affordability
 Finally, I try to remove the year

```{r}


model_gam_females.noyear <- gam(
  Prevalence_females ~
    Country+
    s(HDI, bs = 'cr')+
    Affordability,
  data = data_gam
)
#summary(model_gam_females.noyear)
#plot(model_gam_females.noaffordability)

summ_females<-summary(model_gam_females.noaffordability)
T0.females<-abs(summ_females$p.table[2,3])
T0.females

res.females<-model_gam_females.noyear$residuals
fitted.values.females<-model_gam_females.noyear$fitted.values

T2.females<-numeric(B)
n<-nrow(data_gam)
set.seed(seed)
for(perm in 1:B){
  permutation <- sample(n)
  res.perm <- res.females[permutation]
  response.perm <- fitted.values.females + res.perm
  model.perm<-gam(
    response.perm ~
      data_gam$Year +
      data_gam$Country+
      s(data_gam$HDI, bs = 'cr') +
      data_gam$Affordability
  )
  T2.females[perm] <- abs(summary(model.perm)$p.table[2,3])
}
diagnostic_permutation(T0.females,T2.females)
```

We should keep the year

```{r}
###Final model females
final_model_females<-gam(
  Prevalence_females ~
    Year +
    Country+
    s(HDI, bs = 'cr') +
    Affordability,
   data=data_gam
)
#summary(final_model_females)

plot(final_model_females,  main="HDI smooth term")
```

###################
###################
## Bootstrap on the prediction for the final model for female prevalence
###################
###################
We now compute bootstrap prediction intervals on predictions for 2025, to assess which countries will reach DSG3 goal of relative 30% decrease from 2010 to 2025

```{r}
prevf<-data_gam[data_gam$Country=="France",3][1]
prevf
target<- prevf-0.3*prevf
target

new_obs<-data.frame(Country="France",Year=2025,HDI=0.902,Affordability=3.2)
predict(final_model_females,new_obs,se=TRUE)


xnew = new_obs 
T.obs = predict(final_model_females,newdata=xnew)

fitted.obs<-predict(final_model_females, data_gam) 
res.obs<- data_gam$Prevalence_females-fitted.obs #change y

T.boot <- numeric(B)
set.seed(seed)
pb = progress::progress_bar$new(total = B,
                                format = " Processing [:bar] :percent eta: :eta")
for(i in 1:B){
  response.b <- fitted.obs + sample(res.obs, replace = T)
  model.boot <- gam(response.b ~ 
                      Year +
                      Country+
                      s(HDI, bs = 'cr') +
                      Affordability,
                    data=data_gam)
  T.boot[i] <- predict(model.boot,newdata=xnew)
  pb$tick()
}
myalpha=0.1
diagnostic_bootstrap(T.boot, T.obs, alpha = myalpha)
target
```

France will not reach its SDG3 target for females, according to this prediction

We now try for Turkey 

```{r}
prevf<-data_gam[data_gam$Country=="T端rkiye",3][1]
prevf
target<- prevf-0.3*prevf
target
new_obs<-data.frame(Country="T端rkiye",Year=2025,HDI=0.85,Affordability=3.8)
xnew = new_obs 
T.obs = predict(final_model_females,newdata=xnew)

fitted.obs<-predict(final_model_females, data_gam) 
res.obs<- data_gam$Prevalence_females-fitted.obs #change y

T.boot <- numeric(B)
set.seed(seed)
pb = progress::progress_bar$new(total = B,
                                format = " Processing [:bar] :percent eta: :eta")
for(i in 1:B){
  response.b <- fitted.obs + sample(res.obs, replace = T)
  
  model.boot <- gam(response.b ~ 
                            Year +
                            Country+
                            s(HDI, bs = 'cr') +
                            Affordability,
                          data=data_gam)
  T.boot[i] <- predict(model.boot,newdata=xnew)
  pb$tick()
}
myalpha=0.1
diagnostic_bootstrap(T.boot, T.obs, alpha = myalpha)
target
```

Turkey will not reach its SDG3 target for females, according to this prediction

We now try for Norway

```{r}

prevf<-data_gam[data_gam$Country=="Norway",3][1]
prevf
target<- prevf-0.3*prevf
target

new_obs<-data.frame(Country="Norway",Year=2025,HDI=0.966,Affordability=2.22)
predict(final_model_females,new_obs,se=TRUE)
xnew = new_obs 
T.obs = predict(final_model_females,newdata=xnew)

fitted.obs<-predict(final_model_females, data_gam) 
res.obs<- data_gam$Prevalence_females-fitted.obs 

T.boot <- numeric(B)
set.seed(seed)
pb = progress::progress_bar$new(total = B,
                                format = " Processing [:bar] :percent eta: :eta")
for(i in 1:B){
  response.b <- fitted.obs + sample(res.obs, replace = T)
  model.boot <- gam(response.b ~ 
                      Year +
                      Country+
                      s(HDI, bs = 'cr') +
                      Affordability,
                    data=data_gam)
  T.boot[i] <- predict(model.boot,newdata=xnew)
  pb$tick()
}
myalpha=0.1
diagnostic_bootstrap(T.boot, T.obs, alpha = myalpha)
target
```

Norway will probaby reach the target





################################################################################
################################################################################
#######                           EXTRA1:  QGAM                      ##########
################################################################################
################################################################################

We run regression on quantiles, instead than on the mean. We start from the median

```{r}

final_model_females_qgam<-qgam(
  Prevalence_females ~
    Year +
    Country+
    s(HDI, bs = 'cr') +
    Affordability,
  data=data_gam,
  qu=0.5
  #  try to remove the outliers with respect to HDI- but that does not change much
  # data= data_gam[!data_gam$Country %in% c("Mexico", "Chile", "Colombia","Costa Rica"), ]
)


summary(final_model_females_qgam)
plot(final_model_females_qgam,main="qgam")
#plot(final_model_females_qgam, scale = FALSE, pages = 1)
```

We try to predict the value for Turkey
```{r}
prevf<-data_gam[data_gam$Country=="T端rkiye",3][1]
prevf
target<- prevf-0.3*prevf
target

new_obs<-data.frame(Country="T端rkiye",Year=2025,HDI=0.843,Affordability=3.1)
predict(final_model_females_qgam,new_obs,se=TRUE)

```


For a lack of time, we could not try conformalized prediction intervals for quantile regression, but will try them in the future.

################################################################################
################################################################################
#######                              EXTRA:                           ##########
################################################################################
################################################################################



First using MPE metric defined as follows

```{r}
# Specify the number of folds for cross-validation
num_folds <- 5

# Create an empty vector to store the evaluation metric for each fold
evaluation_metric <- numeric(num_folds)

# Perform cross-validation
set.seed(2022)  # For reproducibility

# Create indices for cross-validation folds
folds <- sample(rep(1:num_folds, length.out = nrow(data_gam)))

mpe <- function(actual, predicted) {
  n <- length(actual)
  mpe_val <- (1/n) * sum(abs((actual - predicted)) / actual) * 100
  return(mpe_val)
}

# Iterate over  each fold
for (i in 1:num_folds) {
  # Split the data into training and validation sets based on the fold
  train_data <- data_gam[folds != i, ]  # Training set
  valid_data <- data_gam[folds == i, ]  # Validation set
  
  # Fit the GAM model with GCV using the training set
  gam_model <- gam(Prevalence_females ~
                     Year +
                     Country+
                     s(HDI, bs = 'cr') +
                    # s(Education_females, bs = 'cr') +
                     Affordability,
                   data = train_data)
  
  # Make predictions on the validation set
  predicted_values <- predict(gam_model, newdata = valid_data)
  
  # Calculate the evaluation metric (e.g., MSE) for the fold
  evaluation_metric[i] <- mpe(valid_data$Prevalence_females, predicted_values)
}
plot(valid_data$Prevalence_males,predicted_values)
# Calculate the average evaluation metric across all folds
average_metric <- (mean(evaluation_metric))

# Print the average evaluation metric
cat("Average Evaluation Metric:", average_metric, "\n")
# 0.0001605886
```

Then using MSE 

```{r}
data_gam$HDI_MHI_clustering<-as.factor(data_gam$HDI_MHI_clustering)
#######
model_gam_both <- gam(
  Prevalence_both ~
    Year +
    s(HDI, bs = 'cr') +
    s(GDP, bs = 'cr') +
    s(Education, bs = 'cr') +  
    HDI_MHI_clustering+
    Affordability,
  data = data_gam
)

summary(model_gam_both)


set.seed(2022)  # For reproducibility

# Create indices for cross-validation folds

data_gam$Prevalence_females<-data_gam$Prevalence_females
num_folds=5
folds <- sample(rep(1:num_folds, length.out = nrow(data_gam)))
evaluation_metric<-numeric(num_folds)
# Iterate over each fold
for (i in 1:num_folds) {
  # Split the data into training and validation sets based on the fold
  train_data <- data_gam[folds != i, ]  # Training set
  valid_data <- data_gam[folds == i, ]  # Validation set

  # Fit the GAM model with GCV using the training set
  gam_model <- gam(Prevalence_females ~
                     Year+
                     Country+
                     s(HDI, bs = 'cr') +
                     Affordability,
                   data = train_data
                   )
  # Make predictions on the validation set
  predicted_values <- predict(gam_model, newdata = valid_data)
  #predicted_values <- plogis(predicted_values_l)
  
  #Calculate the evaluation metric (e.g., MSE) for the fold
  evaluation_metric[i] <- mean((valid_data$Prevalence_females - (predicted_values))^2)
}

# Calculate the average evaluation metric across all folds (e.g. RMSE)
average_metric <- (sqrt(mean(evaluation_metric)))

# Print the average evaluation metric
cat("Average Evaluation Metric:", average_metric, "\n") #1.250058

```

